Chapter 16: Particle-laden gravity currents: the lock-release slumping regime at the laboratory scale

### 2024-02-21 paper/main.tex
Pass doc to single column for working
Adding file for answering referees

\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx, caption}
\usepackage{authblk}
\usepackage{textcomp}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{cuted}

@@

\newcommand*{\cyril}{\textcolor{ForestGreen}}
\newcommand*{\marie}{\textcolor{Cyan}}
\newcommand*{\julien}{\textcolor{Brown}}
\newcommand*{\yvan}{\textcolor{Blue}}

\definecolor{jeanc}{RGB}{253,129,43}
\newcommand*{\jean}{\textcolor{jeanc}}

@@

\date{{\it\color{red}Deadline: May 31$^{st}$ - repoussé (mi juillet ou EuroMech?)}}

@@

\twocolumn[

@@

	\vspace{0.3cm}
	{\it \color{red}REMINDER: Word limit of about 10,000 per article, plus 7 figures.}\\
	\vspace{0.3cm}
]

@@

For a finite volume lock-release device on a mild enough slope, the slumping regime necessarily ends when the rarefaction wave associated with finite volume reaches the current front \citep[e.g.][]{Rottman1983, Hogg2006}. This leads to a new regime usually referred to as the inertial regime in the literature. This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
A simple approach can be useful to provide a rapid estimation of the position $X_{\text t}$ or time $T_{\text t}$ at which this transition occurs. For that purpose, it is worth mentioning that such a transition occurs as a rarefaction wave evolves faster than the front. Then, assuming the front velocity $U_{\text{c}}$ to be different from the rarefaction wave velocity at the interface $\sqrt{\tilde{g}H_0}$, which is assumed to be constant along its entire propagation path for simplicity, this leads to a transition length scale $X_{\text t}$ written as
\begin{equation}
	\label{eq:Xt_inertial}
	\displaystyle X_{\text t}\propto  \frac{2L_0 F}{1 - F}.
\end{equation}
Note that here $2$ comes from the path of propagation of the wave: emanating from the gate, it propagates backwards to the end of the reservoir, reflects on the wall and then propagates towards the front.
During the slumping regime, the previous estimation of $F$ \eqref{eq:Froude_Huppert1980} for full depth release $H_0=H_{\text a}$ leads to $X_{\text t}\propto 2L_0/[2(\rho_{\text a}/\rho_{\text c})^{1/2}-1]$. Note that, although more advanced methods based on single-layer shallow-water models predict a similar result, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}, full depth-release experiments have found much larger transitional positions, i.e $X_{\text t} \approx 10 L_{0}$~\citep{Rottman1983}.

@@

Beyond the transition from the slumping regime to other regimes discussed in the previous section (still relevant for PLGCs although parametrisation of the processes involves more complex mechanisms induced by the fluid-particle interaction), here, a new transition towards the deposition of the particle can emerge. In particular, based on a kinematic approach balancing time scales related to current propagation and particle settling, leads to
%
\begin{equation}
	\label{eq:Xt_settling}
	X_{\text t} \propto H_0\frac{U_{\text{c}}}{V_{\text s}}.
\end{equation}
%
Note that such transition shall disappear for the neutrally-buoyant situation of PLGCs.
However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behaviour similar to that of settling~\cite{Schneider2023}. Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear. Overall, the value of $F$ in \eqref{eq:solUc} during the slumping inertial regime remains to clarify varying the influence of the mechanisms discussed so far.

>>>>

% \documentclass[twocolumn]{article}
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx, caption}
\usepackage{authblk}
\usepackage{textcomp}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{cuted}

@@

\usepackage{setspace}
\doublespacing

@@

@@

@@

@@

For a finite volume lock-release device on a mild enough slope, the slumping regime necessarily ends when the rarefaction wave associated with finite volume reaches the current front \citep[e.g.][]{Rottman1983, Hogg2006}. This leads to a new regime usually referred to as the self-similar regime in the literature. This is observed on the front evolution as $U_{\text{c}}$ is no longer time independent but evolves as $t^{-1/3}$.
A simple approach can be useful to provide a rapid estimation of the position $X_{\text t}$ or time $T_{\text t}$ at which this transition occurs. For that purpose, it is worth mentioning that such a transition occurs as a rarefaction wave evolves faster than the front. Then, assuming the front velocity $U_{\text{c}}$ to be different from the rarefaction wave velocity at the interface $\sqrt{\tilde{g}H_0}$, which is assumed to be constant along its entire propagation path for simplicity, this leads to a transition length scale $X_{\text t}$ written as
\begin{equation}
	\label{eq:Xt_inertial}
	\displaystyle X_{\text t}\propto  \frac{2L_0 F}{1 - F}.
\end{equation}
Note that here $2$ comes from the path of propagation of the wave: emanating from the gate, it propagates backwards to the end of the reservoir, reflects on the wall and then propagates towards the front.
During the slumping regime, the previous estimation of $F$ \eqref{eq:Froude_Huppert1980} for full depth release $H_0=H_{\text a}$ leads to $X_{\text t}\propto 2L_0/[2(\rho_{\text a}/\rho_{\text c})^{1/2}-1]$. Note that, although more advanced methods based on single-layer shallow-water models predict a similar result, $X_{\text t} = 2 L_0 F (1 + F/2)^{1/2}$ \citep{Hogg2006}, full depth-release experiments have found much larger transitional positions, i.e $X_{\text t} \approx 10 L_{0}$~\citep{Rottman1983}.

@@

Beyond the transition from the slumping regime to other regimes discussed in the previous section (still relevant for PLGCs although parametrisation of the processes involves more complex mechanisms induced by the fluid-particle interaction), here, a new transition towards the deposition of the particle can emerge. In particular, based on a kinematic approach balancing time scales related to current propagation and particle settling, leads to
%
\begin{equation}
	\label{eq:Xt_settling}
	X_{\text t} \propto H_0\frac{U_{\text{c}}}{V_{\text s}}.
\end{equation}
%
Note that such transition shall disappear for the neutrally-buoyant situation of PLGCs.
However, it has been observed that mixing at the interface can modify the local concentration of particles at the front, resulting in behaviour similar to that of settling~\cite{Schneider2023}. Again, such an approach only gives a scaling of transition, but the influence of the particle dynamics prior to such transition remains unclear. Overall, the value of $F$ in \eqref{eq:solUc} during the slumping regime remains to clarify varying the influence of the mechanisms discussed so far.
